Iterative Monomorphisation

### 2025-02-19 iterative_monomorphisation.tex
mmore
added links

We implemented iterative monomorphisation in Zipperposition~\cite{vukmirovic-et-al-2021}, a higher order prover written in OCaml. Although Zipperposition is polymorphic, it uses the monomorphic prover E~\cite{e} as a backend. Thanks to our work, E can now be used with polymorphic problems. Moreover, our implementation can be used as a preprocessor to interface with other stand-alone provers.

@@

\item Is iterative monomorphisation more effective than the native polymorphism implemented in polymorphic provers?
\end{enumerate}

@@

The monomorphisation algorithm is parametrised by many bounds. The first part of the evaluation process is to find appropriate values for these bounds. The second part compares the performance of Zipperposition with and without E on polymorphic problems. The third part compares the performance of different provers on polymorphic problems and their monomorphised counterparts.
Our raw evaluation data is available online.%
\footnote{\url{TODO}}

The benchmarks are taken from version 8.2.0 of the TPTP library~\cite{tptp}. The library contains 1765 problems in the TF1 and TH1 grammars, corresponding respectively to first-order and higher-order logic with rank-1 polymorphism. Because Zipperposition does not support reasoning with real numbers, we removed all problems that include them. In total, our benchmark suite contains 1534 polymorphic problems. We chose as a measure of success for a given prover (or prover configuration) the number of problems that could each be solved by the prover in at most 30 seconds with a single thread.
%This corresponds to the default time limit in Sledgehammer.

@@

To evaluate the usefulness of iterative monomorphisation as an alternative to native polymorphism, two competitive higher-order polymorphic provers were run on the 1034 problem suite. We chose Leo-III and Zipperposition. We unfortunately had to exclude Vampire because of parsing issues.

>>>>

We implemented iterative monomorphisation in Zipperposition~\cite{vukmirovic-et-al-2021}, a higher order prover written in OCaml. Although Zipperposition is polymorphic, it uses the monomorphic prover E~\cite{e} as a backend. Thanks to our work, E can now be used with polymorphic problems. Moreover, our implementation can be used as a preprocessor to interface with other stand-alone provers.
Our source code is available online.%
\footnote{\url{https://github.com/nartannt/iterative_monomorphisation}}

@@

\item Is iterative monomorphisation more effective than the native polymorphism implemented in polymorphic provers?
\end{enumerate}
%
Our raw evaluation data is available online.%
\footnote{\url{https://zenodo.org/records/14881532}}

@@

The monomorphisation algorithm is parametrised by many bounds. The first part of the evaluation process is to find appropriate values for these bounds. The second part compares the performance of Zipperposition with and without E on polymorphic problems. The third part compares the performance of different provers on polymorphic problems and their monomorphised counterparts.

The benchmarks are taken from version 8.2.0 of the TPTP library~\cite{tptp}. The library contains 1765 problems in the TF1 and TH1 grammars, corresponding respectively to first order and higher order logic with rank-1 polymorphism. Because Zipperposition does not support reasoning with real numbers, we removed all problems that include them. In total, our benchmark suite contains 1534 polymorphic problems. We chose as a measure of success for a given prover (or prover configuration) the number of problems that could each be solved by the prover in at most 30 seconds with a single thread.
%This corresponds to the default time limit in Sledgehammer.

@@

To evaluate the usefulness of iterative monomorphisation as an alternative to native polymorphism, two competitive higher order polymorphic provers were run on the 1034 problem suite. We chose Leo-III and Zipperposition. We unfortunately had to exclude Vampire because of parsing issues.
